Investigating factors that may influence COVID-19 survival

### 2023-10-21 manuscript.Rmd
Update manuscript.Rmd

---
title: Investigating factors that may influence COVID-19 survival
author: 36112985
bibliography: manuscript/bibliography.bib
output:
  rmarkdown::pdf_document:
    keep_tex: FALSE
    citation_package: "default"
    number_sections: FALSE
    extra_dependencies:
      - booktabs
urlcolor: orange
linkcolor: green
citecolor: red
header-includes:
  - \usepackage{color}
  - \usepackage{caption}
  - \captionsetup[figure]{font={normalsize, color=gray}, width=.8\linewidth}
  - \usepackage{xcolor}
  - \usepackage{titlesec}
  - \usepackage{fancyhdr}
  - \newtheorem{theorem}{Theorem}
  - \newtheorem{proposition}[theorem]{Proposition}
  - \newtheorem{remark}{Remark}
  - \newtheorem{definition}{Definition}
  - \definecolor{darkgrey}{HTML}{4a4a4a}
  - \titleformat*{\section}{\Large\color{darkgrey}}
  - \pagestyle{fancy}
---

<!--- Global Settings
    th -> top or here
--->
```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(knitr.table.format = "latex")
```

<!--- imports --->
```{r include = FALSE}
library(xtable)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)

library(magrittr)
library(latex2exp)
library(survival)
library(patchwork)
```

<!--- external functios --->
```{r}
sys.source(file = 'R/evaluation/KaplanMeier.R', envir = knitr::knit_global())
sys.source(file = 'R/functions/StudyData.R', envir = knitr::knit_global())

sys.source(file = 'R/background/Numbers.R', envir = knitr::knit_global())
```

```{r}
data <- StudyData()
```

> # ABSTRACT
>
> **Objective** *To investigate, amongst a set of probable prognostic factors, which factors may influence coronavirus 19 disease survival probabilities.*
>
> **Design** *Prospective cohort study; ongoing.*
>
> **Setting** *The International Severe Acute Respiratory and Emerging Infections Consortium (ISARIC) World Health Organization (WHO) Clinical Characterisation Protocol UK (CCP-UK) study, which enrols in-patients at at-least 260 England, Scotland, and Wales hospitals; the ISARIC Coronavirus Clinical Characterisation Consortium (ISARIC 4C) conducts the study. The investigation's patients are patients enrolled in a participating England hospital between 10 February 2020 and 5 July 2020.*
>
> **Participants** *Each study enrollee has a confirmed SARS-CoV-2 infection, or a high infection likelihood.  The investigation's patients are members of age groups *30 - 39*, *40 - 49*, and higher.*
>
> **Outcome** Patient survival.
>
> **Results** Inconclusive.
>
> **Conclusions**  Insufficient spectrum of factors.

# Introduction

During the past 2 years, infections due to severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2) have led to a high number of coronavirus 19 disease deaths and in-hospital care.  With time, case fatality rates, and in-hospital survival rates, have improved [...] as our understanding of coronavirus 19 disease pathogenesis has developed.  Consequently,  disease prevention, mitigation, and treatment measures have improved.

Effective disease treatment measures depend on our understanding of the disease's prognosis factors [@riley2013], the focus of prognostic factors research.  Prognostic factor research *aims to identify factors associated with subsequent clinical outcome in people with a particular disease or health condition* [@riley2013].

This brief investigates the prognostic factor potential of a set of measures associated with a prospective cohort study. The study's hypothesis is that *one or more of the measures has a significant impact on each patient's survival probability*.  The ISARIC 4C study outcomes are: discharged alive, remains in hospital, death, transferred, and palliative discharge. Herein, the overarching outcomes of interest are: dead, alive.  The predictors, factors, under investigation

# Methods

@@

This investigation's data is an indirect ISARIC (International Severe Acute Respiratory Infection Consortium)  4C (Coronavirus Clinical Characterisation Consortium) study data set.  The study is an ongoing prospective cohort study based in more than 200 United Kingdom acute care hospitals.  This investigation's data is an England only data set.  The reporting ...

@@

## Participants

Inclusion criteria were people of all ages who were admitted to one of 208 acute care hospitals in England, Scotland, and Wales with proven or high likelihood of infection with a pathogen of public health interest, defined as SARS-CoV-2 for this event by Public Health England. Reverse transcriptase polymerase chain reaction was the only mode of testing available during the period of study. The decision to test was at the discretion of the clinician attending the patient, and not defined by protocol. The enrolment criterion “high likelihood of infection” reflects that a preparedness protocol cannot assume that a diagnostic test will be available for an emergent pathogen. Site training emphasises that only patients who tested positive for covid-19 were eligible for enrolment.

National guidance was provided by Public Health England and other UK public health agencies that advised who to test based on clinical case definitions for possible covid-19. We also included patients who had been admitted for a separate condition but had tested positive for covid-19 during their hospital stay. Patients were only enrolled during their index admission. Patients had clinical information from their routine health records uploaded into the case report form. Consent was not required for collection of depersonalised routine healthcare data for research in England and Wales. A waiver for consent was given by the Public Benefit and Privacy Panel in Scotland.

## Data

The ISARIC 4C study collects its baseline data via paper case report forms developed by ISARIC, and the WHO (World Health Organization), for outbreak investigations.  Additionally, the data brief states that the baseline data records are *uploaded [on] admission, and usually before hospital episodes were complete, to a REDCap database (Research Electronic Data Capture, Vanderbilt University, US, hosted by University of Oxford, UK).*

## Exploratory Data Analysis

* Relationships
* Correlations

@@

* missing completely at random (MCAR) / administrative errors, accidents
* missing at random (MAR) / missing data associated with known patient characteristics [independent variables], or the outcome
* missing not at random (MNAR) / missing data associated with missing values of the factor/predictor in question or with unobserved predictors

@@

Herein, missing data mechanisms analysis is via logistic regression.  Almost all the independent variables have missing values patterns that violate MCAR; addressed via multiple imputation in-line with the advice and protocols of [@sterne2009] & [@steyerberg2010].

## Factor Analysis

The Cox Hazard, Restricted Mean Survival

@@

# Results

## Characteristics

\begin{table}[h!]
    \centering
    \renewcommand{\arraystretch}{1.25}
    \begin{tabular}{l r r r}
    & & & \\
    \textbf{In Hospital Mortality} & & & \\
    \hline
     & \textbf{no} & \textbf{yes} & \textbf{NA} \\
    deceased & 31724 & 14049 & 347 \\
     & 0.688 & 0.305 & 0.008 \\
     & & & \\
    \textbf{Comorbidities} & & & \\
    \hline
     & \textbf{no} & \textbf{yes} & \textbf{NA} \\
    asthma & 37376 & 5872 & 2872 \\
    & 0.81 & 0.13 & 0.06 \\
    mild liver disease & 42021 & 666 & 3433 \\
     & 0.91 & 0.01 & 0.07 \\
    renal disease & 35158 & 8056 & 2906 \\
     & 0.76 & 0.17 & 0.06 \\
    pulmonary & 35265 & 8078 & 2777 \\
     & 0.76 & 0.18 & 0.06 \\
    neurological disorder & 37227 & 5806 & 3087 \\
     & 0.81 & 0.13 & 0.07 \\
    liver mod severe &  41901 &  902 & 3317 \\
     & 0.91 & 0.02 & 0.07 \\
    malignant neoplasm & 38164 & 4734 & 3222 \\
     & 0.83 & 0.10 & 0.07 \\
     & & & \\
    \textbf{Demographics} & & & \\
    \hline
    \textbf{Age Group} &  \textbf{Male} & \textbf{Female} & \textbf{NA} \\
    80-89 & 7067 & 6197 & 31 \\
    & 0.153 & 0.134 & 0.001 \\
    70-79 & 6388 & 4445 & 19 \\
    & 0.139 & 0.096 & 0.000 \\
    60-69 & 4321 & 2601 & 9 \\
    & 0.094 & 0.056 & 0.000 \\
    50-59 & 3463 & 2151 & 9 \\
    & 0.075 & 0.047 & 0.000 \\
    90+ & 2085 & 2816 & 13 \\
    & 0.045 & 0.061 & 0.000 \\
    40-49 & 1846 & 1110 & 5 \\
    & 0.040 & 0.024 & 0.000 \\
    30-39 & 784 & 756 & 4 \\
    & 0.017 & 0.016 & 0.000 \\
    \hline
    \end{tabular}
\end{table}

@@

```{r fig.height = 3.5, fig.align = 'center', warning = FALSE, message = FALSE, fig.cap = "The null survival curves, whereby (A) is based on complete cases only, whilst (B) is based on all plausible records after multiple imputation."}
KaplanMeier()
```

## Null Survival Curve

* Complete Case
* Imputation Case

## Uni-variate & Multivariate Analysis

* The table ... Cox

## Adjusted Analysis via Restricted Mean Survival Time

The proportional hazard assumption does not hold ...

# Discussion

@@

**Proportional Hazards**: The curves.

@@

The investigation opted for the latter, especially because ...

@@

Possible bias points

* By virtue of the study's design pre-study--admission differences between hospitals is possible due to the ambiguity of the study's patient recruitment terms, and this might lead to post-study differences.  In a nutshell, the study is susceptible to selection bias. [Trochim]

* Missing data [@sterne2009]

# Conclusion

The analysis of missing values hints at flaws in data collection.

@@

```{r}

```

Graphs

@@

Clusters, patterns, associations.

<!--- # References {#references .unnumbered} --->
# References

>>>>

---
title: Investigating factors that may influence COVID-19 survival
author: 36112985
bibliography: manuscript/bibliography.bib
output:
  rmarkdown::pdf_document:
    keep_tex: FALSE
    citation_package: "default"
    number_sections: FALSE
    extra_dependencies:
      - booktabs
urlcolor: orange
linkcolor: green
citecolor: red
header-includes:
  - \usepackage{color}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \captionsetup[figure]{font={normalsize, color=gray}, width=.8\linewidth}
  - \usepackage{xcolor}
  - \usepackage{titlesec}
  - \usepackage{fancyhdr}
  - \newtheorem{theorem}{Theorem}
  - \newtheorem{proposition}[theorem]{Proposition}
  - \newtheorem{remark}{Remark}
  - \newtheorem{definition}{Definition}
  - \definecolor{darkgrey}{HTML}{4a4a4a}
  - \titleformat*{\section}{\Large\color{darkgrey}}
  - \pagestyle{fancy}
---

<!--- Global Settings
    th -> top or here
--->
```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(knitr.table.format = 'latex')
options(knitr.kable.NA = '')
```

<!--- imports --->
```{r include = FALSE}
library(xtable)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(matrixStats)

library(magrittr)
library(latex2exp)
library(survival)
library(survRM2)
library(patchwork)
library(kableExtra)
```

<!--- external functios --->
```{r}
sys.source(file = 'R/functions/ExtensiveStudyData.R', envir = knitr::knit_global())

sys.source(file = 'R/background/Numbers.R', envir = knitr::knit_global())
sys.source(file = 'R/background/CorrelationOfPredictors.R', envir = knitr::knit_global())

sys.source(file = 'R/missing/Pattern.R', envir = knitr::knit_global())
sys.source(file = 'R/missing/Test.R', envir = knitr::knit_global())

sys.source(file = 'R/demographics/Is.R', envir = knitr::knit_global())

sys.source(file = 'R/modelling/UnivariateAnalysis.R', envir = knitr::knit_global())
sys.source(file = 'R/modelling/MultivariateAnalysis.R', envir = knitr::knit_global())

sys.source(file = 'R/evaluation/KaplanMeier.R', envir = knitr::knit_global())

sys.source(file = 'R/restricted/RMST.R', envir = knitr::knit_global())
```

```{r}
dataframes <- ExtensiveStudyData(upload = TRUE)
data <- dataframes$data
training_ <- dataframes$training_
data_ <- dataframes$data_
```

> # ABSTRACT
>
> **Objective** *To investigate, amongst a set of probable prognostic factors, which factors may influence coronavirus 19 disease survival probabilities.*
>
> **Design & Setting** *The underlying study is the International Severe Acute Respiratory and Emerging Infections Consortium (ISARIC) World Health Organization (WHO) Clinical Characterisation Protocol UK (CCP-UK) study, which enrols in-patients at at-least 260 England, Scotland, and Wales hospitals; the ISARIC Coronavirus Clinical Characterisation Consortium (ISARIC 4C) conducts the study.  It is an ongoing prospective cohort study. Herein, the patients are patients enrolled in a participating England hospital between 10 February 2020 and 5 July 2020.*
>
> **Results** The results of the investigation are inconclusive. Modelling and analysis could not proceed via the Cox Proportional Hazards model due to the violation of the proportionality assumption.  The restricted mean survival time analysis suggest that sex, malignant neoplasm, neurological disorder, pulmonary disease, and renal disease might have a significant impact on in-hospital coronavirus 19 disease survival.
>
> **Conclusions**  The investigation should be repeated with much more appropriate data, e.g., continuous age variables, start & stop times of events, follow-ups, etc.  The analysis of missing values hints at flaws in data collection, this might be partly addressed via a clearer recruitment protocol.

\vspace{35pt}

# INTRODUCTION

During the past 2 years, infections due to severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2) have led to a high number of coronavirus 19 disease deaths and in-hospital care [@jhu2020, @ritchie2020].  With time, in-hospital survival rates, have improved [@docherty2021].  A number of factors underlie the improvements, one of the most crucial being a better understanding of coronavirus 19 disease, which has led to much more effective treatment measures [@docherty2021].

Effective disease treatment measures partly depend on our understanding of the disease's prognosis factors [@riley2013], the focus of prognostic factors research.  Prognostic factor research *aims to identify factors associated with subsequent clinical outcome in people with a particular disease or health condition* [@riley2013].

The definitive set of factors that affect in-hospital survival probabilities of coronavirus 19 disease patients is unknown, and is a research focus [@gorog2022, @mcgurnaghan2021]. This brief investigation investigates the prognostic factor potential of a set of measures associated with the ISARIC prospective cohort study. The hypothesis being *one or more of the measures has a significant impact on each patient's survival probability*.

\vspace{35pt}

# METHODS

\vspace{10pt}

@@

This investigation's data is an indirect ISARIC (International Severe Acute Respiratory Infection Consortium) 4C (Coronavirus Clinical Characterisation Consortium) study data set; the data set is a project assigned data set, not actively collected.  The ISARIC 4C study is an ongoing prospective cohort study that is conducted by the ISARIC Coronavirus Clinical Characterisation Consortium (ISARIC 4C).  The study enrols in-patients at at-least 260 England, Scotland, and Wales acute care hospitals.  Herein, the patients are patients enrolled in a participating England hospital between 10 February 2020 and 5 July 2020, i.e., this paper's investigation focuses on an England only data set.  Additionally, the England only data set consists of patients within age groups *30 - 39*, *40 - 49*, and higher.*

According to the project brief, each study enrollee has a confirmed SARS-CoV-2 infection, or a high infection likelihood.  The study collects its baseline data via paper case report forms developed by ISARIC, and the WHO (World Health Organization), for outbreak investigations.

@@

## Outcomes

The ISARIC 4C study outcomes are: discharged alive, remains in hospital, death, transferred, and palliative discharge. Herein, the overarching outcomes of interest are: dead, alive.  Hence, investigation's censoring logic is

> Right Censored:
>
>  * discharged alive
>
>  * remains in hospital
>
>  * transferred
>
>  * palliative discharge

> Uncensored:
>
>  * death

Without more information it will be ill-advised to assume that death is the final outcome of patients released under palliative discharge.  Especially because in addition to *end of life care*, palliative care also includes *supportive care* for curable diseases, i.e., support for patients that are expected to recover, but who need critical support whilst convalescing. [@apm2019]

@@

\vspace{10pt}

* Missing Completely at Random (MCAR): administrative errors, accident.

* Missing at Random (MAR): missing data associated with known patient characteristics [independent variables], or the outcome.

* Missing Not at Random (MNAR): missing data associated with missing values of the factor/predictor in question or with unobserved predictors.

\vspace{10pt}

@@

The missing data mechanisms analysis [Appendix] suggests that almost all the independent variables have missing values patterns that violate MCAR.  Consequently, a multiple imputation step, in-line with the advice and protocols of [@sterne2009] & [@steyerberg2010], is part of this investigation.

@@

## Modelling & Analysis: Potential Prognostic Factors

The potential, candidate, prognostic factors are sex, age group, asthma, mild liver disease, moderate/severe liver disease, renal disease, chronic pulmonary disease, chronic neurological disorder, and malignant neoplasm.  Their effects, or otherwise, on survival probabilities, are explored via the Cox Proportional Hazard Model [@cox1972] and the Restricted Mean Survival Time [@royston2013].

\pagebreak

# RESULTS

\vspace{10pt}

## Descriptive Statistics

The numbers of *Table 1* summarise the death, comorbidity , and demographic numbers of the investigation's data set. Of the received data set of $50,000$ records, $3,880$ are considered implausible because each has an outcome date that precedes its admission date.  Consequently, each outlined modelling and analysis step is due to $46,120$ records.

\footnotesize
```{r}
special <- SpecialNumbers(data = data)

special %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    booktabs = TRUE,
    digits = 2,
    escape = FALSE,
    row.names = FALSE,
    col.names = kableExtra::linebreak(c('', 'yes', 'no', 'unknown', 'yes', 'no', 'unknown' )),
    caption = 'Data Description')  %>%
  kableExtra::kable_styling(position = 'center') %>%
  kableExtra::add_header_above(c('', 'N' = 3, 'Percentage' = 3)) %>%
  kableExtra::column_spec(column = 1, latex_column_spec = "p{0.17\\\\linewidth}l")

```
\normalsize

\pagebreak

The graph of *fig. 1* illustrates the degree of correlation between the potential prognostic factors.  The embedded Cramer's V values suggest weak associations only; the degree of freedom per cell is one.

```{r out.width = '65%', fig.align = 'center', results = 'hide', warning = FALSE, message = FALSE, fig.cap = 'Is there a degree of correlation between predictors?'}
CorrelationOfPredictors(
  predictors = c('age_group', 'sex', 'asthma', 'liver_mild', 'renal', 'pulmonary',
                 'neurological', 'liver_mod_severe', 'malignant_neoplasm'),
  data = data)
```

@@

The graphs of *fig 2* illustrate the survival proportions over time.  The probability of survival drops sharply during the first 200 days, this unsteady pattern suggest that the risk of death over time might not be constant.

```{r fig.height = 3.5, out.width = '85%', fig.align = 'center', warning = FALSE, message = FALSE, fig.cap = "The null survival curves, whereby (A) is based on complete cases only, whilst (B) is based on all plausible records after multiple imputation."}
KaplanMeier()
```

\pagebreak

## Univariate Analysis: Cox Proportional Hazards Model

This Cox Proportional Hazard Model uni-variate analysis models & analyses the separate effect of each potential prognostic factor on survival probability.  The analysis is via the imputed data set, which excludes the *3,880* implausible observation, i.e., there are *46,120* records; *50,000* - *3,880*.

In terms of age groups, the reference age group is *80 – 89*, and their $p_{value}$ scores suggests that **(a)** each age group, except *90+*, has a lower risk of death compared to age group *80 – 89*, and **(b)** the risk of death increases with age.  The sex $p_{value}$ suggests that females have a lower risk of death compared to males.  The $p_{value}$ scores for patients with an asthma, renal disease, pulmonary disease, or malignant neoplasm, diagnosis suggests that these patients have a higher risk of death compared to their respective disease/disorder free counterparts.

Alas, the proportionality tests invalidate most of the preceding observations; age group, malignant neoplasm, and neurological disorder have a score test value below $0.05$, i.e., the proportionality assumption does not hold.  The multivariate analysis confirms this observation.

\footnotesize
```{r}
diagnostics <- UnivariateAnalysisData()

diagnostics %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    booktabs = TRUE,
    digits = 3,
    escape = FALSE,
    col.names = kableExtra::linebreak(c('', 'coefficient', 'hazard\nratio', 'lower',
                                        'upper', '$p_{value}$', '${U}$' )),
    caption = 'Univariate Analysis: Imputed Data', position = '!ht') %>%
  kableExtra::kable_styling(position = 'center') %>%
  kableExtra::add_header_above(c('', '', '', 'hazard ratio\nCI' = 2, '', '')) %>%
  kableExtra::add_footnote(c('CI: Confidence Interval',
                             'U: Proportionality score test'), notation = "number")  %>%
  kableExtra::column_spec(column = 1, latex_column_spec = "p{0.20\\\\linewidth}l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "r")
```
\normalsize

\pagebreak

## Multivariate Analysis: Cox Proportional Hazards Model

Similar to the univariate analysis, the proportionality tests invalidate the multivariate analysis results of *table 3*; age group, malignant neoplasm, and neurological disorder have score test values below $0.05$.

\footnotesize
```{r}
diagnostics <- MultivariateAnalysisData()

diagnostics %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    booktabs = TRUE,
    digits = 3,
    escape = FALSE,
    col.names = kableExtra::linebreak(c('', 'coefficient', 'hazard\nratio', 'lower',
                                        'upper', '$p_{value}$', '${U}$' )),
    caption = 'Multivariate Analysis: Imputed Data', position = '!ht') %>%
  kableExtra::kable_styling(position = 'center') %>%
  kableExtra::add_header_above(c('', '', '', 'hazard ratio\nCI' = 2, '', '')) %>%
  kableExtra::add_footnote(c('CI: Confidence Interval',
                             'U: Proportionality score test'), notation = "number")  %>%
  kableExtra::column_spec(column = 1, latex_column_spec = "p{0.20\\\\linewidth}l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "r")
```
\normalsize

\pagebreak

## Unadjusted & Adjusted Analysis: Restricted Mean Survival Time

The restricted mean survival time (RMST) is an alternative survival analysis approach, especially when the risk over time might be inconstant and/or the Cox Proportional Hazards Model assumptions do not hold. The restricted mean survival time is *... a measure of average survival from time 0 to a specified time point ...* [@royston2013]  The results of *Table ...* summarise the unadjusted and adjusted uni-variate RMST analysis of the named predictors.

In the case of the adjusted analysis, a continuous variable field of age group medians, instead of the age group field, is used; access to the original age values would have been much better.  As usual, the adjusted analysis is the case whereby a single predictor is in focus whilst controlling for other variables/predictors, and the hypotheses per predictor are

\vspace{15pt}

* $H_{0}: \tau_{yes} = \tau_{no}$
* $H_{1}: \tau_{yes} \neq \tau_{no}$

\vspace{15pt}

Wherein the null hypothesis is that the survival time w.r.t. (with respect to) yes responses is equivalent to the survival time w.r.t. no responses.  Noting that the specified time point is 267 days $\longrightarrow$ minimum of the *maximum time to outcome* across all predictors

\vspace{15pt}

```r
  # for a single predictor
  min(max(data[data[, predictor] == 1, 'time_to_outcome']),
      max(data[data[, predictor] == 0, 'time_to_outcome']))
```

\vspace{15pt}

Then the **adjusted analysis results** of *table 4* suggests that within a 267 days limit,

* on average female patients survived 18.46 days longer than males.
* the survival days of patients with malignant neoplasm is approximately 18 days shorter than those without.
* on average patients with a neurological disorder had approximately 31 fewer survival days.
* on average patients with a pulmonary disease had approximately 12 fewer survival days.
* on average patients with a renal had approximately 13 fewer survival days.

There are no other significant results.  Further analysis, much more appropriate data, e.g., continuous age values, and a wider range of potential factors [@docherty2021], is required in order to make definitive conclusions.

\footnotesize
```{r}
diagnostics <- RMST()

diagnostics %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    booktabs = TRUE,
    digits = 2,
    escape = FALSE,
    col.names = kableExtra::linebreak(c('', 'est.', 'lower', 'upper', '$p_{value}$',
                                        'est.', 'lower', 'upper', '$p_{value}$')),
    caption = 'Restricted Mean Survival Time Modelling and Analysis', position = '!ht') %>%
    kableExtra::kable_styling(position = 'center') %>%
    kableExtra::add_header_above(c('', '', 'CI' = 2, '', '', 'CI' = 2, '')) %>%
    kableExtra::add_header_above(c('', 'Unadjusted' = 4, 'Adjusted' = 4)) %>%
    kableExtra::column_spec(column = 1, latex_column_spec = "p{0.20\\\\linewidth}l")
```
\normalsize

\vspace{65pt}

\pagebreak

# DISCUSSION

@@

**Proportional Hazards**: The proportionality tests illustrated that the data violates the proportionality assumption.

@@

The investigation opted for the latter, due to the missing data patterns observed (Appendix)

@@

There are a few possible bias points.  By virtue of the study's design pre-study--admission differences between hospitals is possible due to the ambiguity of the study's patient recruitment terms, and this might lead to post-study differences.  In a nutshell, the study is susceptible to selection bias. [Trochim]  Bias due to missing data [@sterne2009].

## NHS England COVID-19 Hospital Activity & ISARIC Study Excerpt

It is probably impossible to compare the study's patient population to the NHS England patient population of the same
time period.  There are a few reasons why $\longrightarrow$ the COVID-19 Hospital Activity Data:

* Does not include a daily record of new COVID-19 patients admissions.  Instead, it continuously updates a number
that denotes *the number of admitted COVID-19 patients that were still in hospital during the last 24 hours*
* Does not include in-hospital COVID-19 deaths.
* Has a different age groups demarcation: `[0  5]`, `[6  17]`, `[18  64]`, `[65  84]`, `85+`

Hence, even if the results had been conclusive, inference to the population might have been tricky.

\pagebreak

# CONCLUSION

The results of the investigation are inconclusive. Modelling and analysis could not proceed via the Cox Proportional Hazards model due to the violation of the proportionality assumption.  The restricted mean survival time analysis suggest that sex, malignant neoplasm, neurological disorder, pulmonary disease, and renal disease might have a significant impact on in-hospital coronavirus 19 disease survival.   The analysis of missing values hints at flaws in data collection.

\pagebreak

@@

The study's female/male population is not representative of England's population, therefore population inference is not appropriate.

```{r fig.height = 3.5, out.width = '49%', fig.align = 'center', warning = FALSE, message = FALSE, fig.cap = "Is the investigation's population representative of England's population?"}
Is()
```

\vspace{10pt}

@@

The clusters, patterns, associations, in relation to missing values.

\vspace{10pt}

```{r out.width = '65%', fig.align = 'center', fig.cap = 'Fractions denoting the instance proportions whereby a pair of predictors are both NA values'}
PatternDendogramNA(data = data)
```

\vspace{10pt}

```{r out.width = '65%', fig.align = 'center', fig.cap = 'The proportion of missing values per predictor variable'}
PatternVariableNA(data = data)
```

\vspace{10pt}

```{r warning = FALSE, message = FALSE, include = FALSE}
diagnostics <- Test(data = data)

diagnostics %>%
  dplyr::mutate_all(linebreak) %>%
    kableExtra::kable(
      format = 'latex',
      booktabs = TRUE,
      digits = 2,
      escape = FALSE,
      row.names = FALSE,
      col.names = kableExtra::linebreak(c('Missing Value\nCells of', 'Associated with one or more\naux./pred. variables' )),
      caption = 'Data Description') %>%
    kableExtra::column_spec(column = 1, latex_column_spec = "p{0.3\\\\linewidth}l")
```

\pagebreak
\pagebreak
<!--- # References {#references .unnumbered} --->
# References
